\ttitle

### 2013-11-01 Appendices/AppendixA-summaries.tex
Prefix refs with non-breaking space and capitalize refs

\label{AppendixA} % For referencing this appendix elsewhere, use \ref{AppendixA}

@@

Linear models are well studied, in particular for the problem of linear regression.
This can be solved by the \emph{least-squares} method and classification as discussed in section \ref{least-squares}, the \emph{perceptron} as explained in section \ref{perceptron}.
Linear regression with the \emph{support vector machine} is handled in section \ref{svm-explained} and used for probability density estimation in section \ref{linear-classifier-pdf}.
The kernel trick used for learning non-linear models is explained in section \ref{non-linear}.

@@

When the algorithm is completed, every $y_i\vectorsym{x}_i$ is added $\alpha_i$ times to the weight vector (every time it was misclassified).
Thus, the weight vector can be expressed as: $\vectorsym{w} = \sum_{i=i}^n \alpha_i y_i \vectorsym{x}_i$.
In other words: the weight vector is a linear combination of the training instances.
The dual form of the algorithm learns the instance weights $\alpha_i$ rather than the features weights $\vectorsym{w}_i$.
An instance $\vectorsym{x}$ is then classified as $\hat{y} = sign(\sum_{i=1}^n \alpha_i y_i \vectorsym{x}_i \cdot \vectorsym{x})$.
This means that during the training only the pairwise dot-products of the data is needed; this results in the \emph{n}-by-\emph{n} Gram matrix $\matrixsym{G} = \matrixsym{X}\matrixsym{X}^T$.
This instance-based perspective will be further discussed in section \ref{svm-explained} about the support vector machine.

@@

This equation can be transformed with the Lagrange multipliers by adding the constraints to the minimization part with multipliers $\alpha_i$.
Taking the partial derivative with respect to $t$ and setting it to $0$, we find that for the optimal solution (threshold) $t$ we have $\sum_{i=1}^{n} \alpha_i y_i = 0$.
When we take the partial derivative with respect to $w\vectorsym{w}$ we see that the Lagrange multipliers define the weight vector as a linear combination of the training examples.
This partial derivative is $0$ for an optimal weight we get that $\vectorsym{w} = \sum_{i=1}^{n} \alpha_i y_i \vectorsym{x}_i$, \emph{which is the same expression as for the perceptron derived in section \ref{perceptron}}.
By plugging $\vectorsym{w}$ and $t$ back into the Lagrange equation, we can eliminate these and get the dual optimization problem entirely formulated in terms of the Lagrange multipliers:

@@

The second property enables powerful adaption for support vector machines to learn non-linear decision boundaries, as discussed in section \ref{non-linear}.

@@

\subsection{Change Point Detection In Time Series Data Using Support Vectors}
Paper by Fatih Camci \cite{camci2010change}
About segmentation with SVMs.
Will be main material for section \ref{sec:svm_in_change_detection} about SVMs.

@@

\subsubsection{Problem formulation}
[Not summarized here, useful for e.g. section \ref{statistical-framework}]

\subsubsection{SVCPD: The algorithm}
Instead of using statistical properties of the data, for each window of size $w$ a hyper-sphere is constructed without increasing computational complexity due to the \emph{kernel trick}.
The window size is related to the sensitivity of the method to change; small windows are sensitive with high false alarm rate whilst large windows are slow to detect change and have low alarm rates.
The algorithm is listed in table \ref{tab:alg-svcdp}.
Note that SVCPD can be applied directly to multidimensional data, whilst many other methods can only be applied to one-dimensional data.

@@

The (expected) value of $D_k$ given a change in variance differs in the context.
If a smaller variance corresponds to the smaller portion of the series, then it will be harden to find the change point using $D_k$.
There is a masking effect when there are multiple change points in the series; the order of small, medium and large variances result in the value of $D_k$.
The iterative algorithm presented in this paper in section \ref{subsec:icss} is designed to lessen the masking effect.

%--------------------------------------------
\subsubsection{Multiple changes: Iterated cumulative sums of squares}\label{subsec:icss}
In case of a single change point the $D_k$ method would succeed.
But we are interested in multiple change points of variance, and thus the usefulness of the $D_k$ reduces due to the masking effect.
A solution is to iteratively applying the method and dividing the series at each possible change point.
The algorithm is presented in table \ref{tab:alg-icss}.
It is the third steps which reduces the masking effect and helps to ``fine tune'' the algorithm by (re)moving the potential change points by checking each location given the adjacent ones.

>>>>

\label{AppendixA} % For referencing this appendix elsewhere, use~\ref{AppendixA}

@@

Linear models are well studied, in particular for the problem of linear regression.
This can be solved by the \emph{least-squares} method and classification as discussed in Section~\ref{least-squares}, the \emph{perceptron} as explained in Section~\ref{perceptron}.
Linear regression with the \emph{support vector machine} is handled in Section~\ref{svm-explained} and used for probability density estimation in Section~\ref{linear-classifier-pdf}.
The kernel trick used for learning non-linear models is explained in Section~\ref{non-linear}.

@@

When the algorithm is completed, every $y_i\vectorsym{x}_i$ is added $\alpha_i$ times to the weight vector (every time it was misclassified).
Thus, the weight vector can be expressed as: $\vectorsym{w} = \sum_{i=i}^n \alpha_i y_i \vectorsym{x}_i$.
In other words: the weight vector is a linear combination of the training instances.
The dual form of the algorithm learns the instance weights $\alpha_i$ rather than the features weights $\vectorsym{w}_i$.
An instance $\vectorsym{x}$ is then classified as $\hat{y} = sign(\sum_{i=1}^n \alpha_i y_i \vectorsym{x}_i \cdot \vectorsym{x})$.
This means that during the training only the pairwise dot-products of the data is needed; this results in the \emph{n}-by-\emph{n} Gram matrix $\matrixsym{G} = \matrixsym{X}\matrixsym{X}^T$.
This instance-based perspective will be further discussed in Section~\ref{svm-explained} about the support vector machine.

@@

This equation can be transformed with the Lagrange multipliers by adding the constraints to the minimization part with multipliers $\alpha_i$.
Taking the partial derivative with respect to $t$ and setting it to $0$, we find that for the optimal solution (threshold) $t$ we have $\sum_{i=1}^{n} \alpha_i y_i = 0$.
When we take the partial derivative with respect to $w\vectorsym{w}$ we see that the Lagrange multipliers define the weight vector as a linear combination of the training examples.
This partial derivative is $0$ for an optimal weight we get that $\vectorsym{w} = \sum_{i=1}^{n} \alpha_i y_i \vectorsym{x}_i$, \emph{which is the same expression as for the perceptron derived in Section~\ref{perceptron}}.
By plugging $\vectorsym{w}$ and $t$ back into the Lagrange equation, we can eliminate these and get the dual optimization problem entirely formulated in terms of the Lagrange multipliers:

@@

The second property enables powerful adaption for support vector machines to learn non-linear decision boundaries, as discussed in Section~\ref{non-linear}.

@@

\subsection{Change Point Detection In Time Series Data Using Support Vectors}
Paper by Fatih Camci \cite{camci2010change}
About segmentation with SVMs.
Will be main material for Section~\ref{sec:svm_in_change_detection} about SVMs.

@@

\subsubsection{Problem formulation}
[Not summarized here, useful for e.g. Section~\ref{statistical-framework}]

\subsubsection{SVCPD: The algorithm}
Instead of using statistical properties of the data, for each window of size $w$ a hyper-sphere is constructed without increasing computational complexity due to the \emph{kernel trick}.
The window size is related to the sensitivity of the method to change; small windows are sensitive with high false alarm rate whilst large windows are slow to detect change and have low alarm rates.
The algorithm is listed in table~\ref{tab:alg-svcdp}.
Note that SVCPD can be applied directly to multidimensional data, whilst many other methods can only be applied to one-dimensional data.

@@

The (expected) value of $D_k$ given a change in variance differs in the context.
If a smaller variance corresponds to the smaller portion of the series, then it will be harden to find the change point using $D_k$.
There is a masking effect when there are multiple change points in the series; the order of small, medium and large variances result in the value of $D_k$.
The iterative algorithm presented in this paper in Section~\ref{subsec:icss} is designed to lessen the masking effect.

%--------------------------------------------
\subsubsection{Multiple changes: Iterated cumulative sums of squares}\label{subsec:icss}
In case of a single change point the $D_k$ method would succeed.
But we are interested in multiple change points of variance, and thus the usefulness of the $D_k$ reduces due to the masking effect.
A solution is to iteratively applying the method and dividing the series at each possible change point.
The algorithm is presented in table~\ref{tab:alg-icss}.
It is the third steps which reduces the masking effect and helps to ``fine tune'' the algorithm by (re)moving the potential change points by checking each location given the adjacent ones.

### 2013-11-01 Appendices/AppendixC.tex
working on outline of proposed method
WIP
Prefix refs with non-breaking space and capitalize refs

%---------------------------------------------
%---------------------------------------------
\clearpage
\section{Joint segmentation of multivariate time series with hidden process regression for human activity recognition, by Chamroukhi}\label{sec:appendix-C-joint-segmentation}
Paper: \cite{chamroukhi2013joint}.
This approach models the time series data by a parameterized regression, filtered with a \gls{hmm} to smooth out high frequency activity transitions.
With each observation $i$, generated by a $K$-state hidden process, an activity label $z_i$ (and thus sequence) is associated.
Observations follow a regression model:
%
\begin{equation}
  y_i = \beta_{z_i}^T \vectorsym{t}_i + \sigma_{z_i} \epsilon_i; \quad \epsilon_i \sim \mathcal{N}(0,1), \quad (i=1, \dots ,n)
\end{equation}
%
The observations get a label assigned by maximizing the logistic probability ($\pi_k$):
%
\begin{equation}
  \hat{z}_i = \argmaxold \limits_{1 \leq k \leq K} \pi_k (t_i; \hat{\vectorsym{w}}), \quad (i=1, \dots, n)
\end{equation}
%

@@

Method and formulas very similair to description in Section \ref{sec:camci}.

%---------------------------------------------
%---------------------------------------------
\clearpage
\section{Support Vector Density Estimation, by Weston et al.}
Paper: \cite{weston1999support}.
Using the notation of \cite{weston1999support}, the distribution function of a density function $p(x)$ is represented as:
%
\begin{equation}
  F(x) = P(X \leq x) = \int_{- \infty}^x p(t)\,\mathrm{d}t
\end{equation}
%
To find the density the following linear equation need to be solved:
%
\begin{equation}
  \int_{-\infty}^\infty \theta(x-t)p(t)\,\mathrm{d}t = F(x)
\end{equation}
%
where
%
\[
  \theta(x) =
  \begin{cases}
    1,& x > 0 \\
    0,& otherwise
  \end{cases}
\]
%
In this problem the distribution function $F(x)$ is unknown and instead we are given the \gls{idd} data $x_1,\dots,x_l$ generated by $F$.

>>>>

%---------------------------------------------
%---------------------------------------------
\clearpage

\section{Joint segmentation of multivariate time series with hidden process regression for human activity recognition, by Chamroukhi}\label{sec:appendix-C-joint-segmentation}
Paper: \cite{chamroukhi2013joint}.
This approach models the time series data by a parameterized regression, filtered with a \gls{hmm} to smooth out high frequency activity transitions.
With each observation $i$, generated by a $K$-state hidden process, an activity label $z_i$ (and thus sequence) is associated.
Observations follow a regression model:
%
\begin{equation}
  y_i = \beta_{z_i}^T \vectorsym{t}_i + \sigma_{z_i} \epsilon_i; \quad \epsilon_i \sim \mathcal{N}(0,1), \quad (i=1, \dots ,n)
\end{equation}
%
The observations get a label assigned by maximizing the logistic probability ($\pi_k$):
%
\begin{equation}
  \hat{z}_i = \argmaxold \limits_{1 \leq k \leq K} \pi_k (t_i; \hat{\vectorsym{w}}), \quad (i=1, \dots, n)
\end{equation}
%

@@

Method and formulas very similair to description in Section~\ref{sec:camci}.

%---------------------------------------------
%---------------------------------------------
\clearpage
\section{Support Vector Density Estimation, by Weston et al.}
Paper: \cite{weston1999support}.
Using the notation of \cite{weston1999support}, the distribution function of a density function $p(x)$ is represented as:
%
\begin{equation}
  F(x) = P(X \leq x) = \int_{- \infty}^x p(t)\,\mathrm{d}t
\end{equation}
%
To find the density the following linear equation need to be solved:
%
\begin{equation}
  \int_{-\infty}^\infty \theta(x-t)p(t)\,\mathrm{d}t = F(x)
\end{equation}
%
where
%
\[
  \theta(x) =
  \begin{cases}
    1,& x > 0 \\
    0,& otherwise
  \end{cases}
\]
%
In this problem the distribution function $F(x)$ is unknown and instead we are given the \gls{iid} data $x_1,\dots,x_l$ generated by $F$.

### 2013-11-01 Chapters/Chapter1.tex
Rough finish of chapter 2
Fix typos, add comments
WIP
Prefix refs with non-breaking space and capitalize refs

\label{Chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1}

@@

\section{Outline}
\begin{itemize}
  \item Context of research (human activity recognition), real-world applications
  \item Current methods, wrapper vs. filter methods
  \item Problem statement with current filter methods (which follows from Chapter \ref{Chapter3} which goes in-depth with methods).
  \item Purpose of this research. E.g. "Find a better algorithm for short-activity segmentation"
  \item Relate to real-world applications
\end{itemize}

>>>>

\label{Chapter1} % For referencing the chapter elsewhere, use~\ref{Chapter1}

@@

\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Context of research (\gls{har}), real-world applications
  \item Current methods, wrapper vs. filter methods
  \item Problem statement with current filter methods (which follows from Chapter~\ref{Chapter3} which goes in-depth with methods).
  \item Purpose of this research. E.g. "Find a better algorithm for short-activity segmentation"
  \item Provide the \emph{motivation} for this approach
  \item Relate to real-world applications
  \item Outline for rest of the thesis
\end{itemize}

-- Overall todo's --
\TODO{The term \emph{emperical data/distruction} is nowhere used in the thesis. Is missing.}

### 2013-11-01 Chapters/Chapter2.tex
Prefix refs with non-breaking space and capitalize refs

\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter2}

@@

The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
It originates from control methods for detection from bench marks.
This method and some derivatiges are discussed and analyzed in section \ref{cusum}.

Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \gls{pdf}, instead of estimating each \gls{pdf}.
This approach is discussed and analyzed in section \ref{density-ratio}.

The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
The method of Camci \cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
A change in the \gls{pdf} is represented by a change in the radius of this sphere.
Section \ref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
\TODO{Put emphasis that this is the source of inspiration for the chosen method?}

% The final method under consideration, change detection via (intrinsic) dimensionality reduction, follows a different point of view.
% Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
% It uses the estimated underlying number of parameters of the time series as a model for change detection.
% Section \ref{dim-reduction} discuses this method.

>>>>

\label{Chapter2} % For referencing the chapter elsewhere, use~\ref{Chapter2}

@@

The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
It originates from control methods for detection from benchmarks.
This method and some derivatiges are discussed and analyzed in Section~\ref{cusum}.

Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf}, instead of estimating each \gls{pdf}.
This approach is discussed and analyzed in Section~\ref{density-ratio}.

The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
The method of Camci \cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
A change in the \gls{pdf} is represented by a change in the radius of this sphere.
Section~\ref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
\TODO{Put emphasis that this is the source of inspiration for the chosen method?}

% The final method under consideration, change detection via (intrinsic) dimensionality reduction, follows a different point of view.
% Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
% It uses the estimated underlying number of parameters of the time series as a model for change detection.
% Section~\ref{dim-reduction} discuses this method.

### 2013-11-01 Chapters/Chapter2/svm.tex
Added the TODO command
Draft, sent to Anne
Prefix refs with non-breaking space and capitalize refs

Many proposals in the field of remote sensing and \gls{har} make use of \glspl{svm} as a model construction method.
An elaborate overview of applications is given in \cite{mountrakis2011support}.
In this section we review methods of change detection based on the applications of \glspl{svm} as model construction method.
A number of proposals have been made using this method.
Sch{\"o}lkopf \etal \cite{scholkopf1999support} applies Vapnik's principle, to never solve a problem which is more general than the one that one is actually interested in, to novelty detection.
In the case of novelty detection, they argue there is no need for density estimation of the distribution.
Simple algorithms estimate the density by considering how many data points fall in a region of interest.
The \gls{nu-svm} method instead starts with the number of data points that should be within the region and estimates a region with that desired property.
It builds on the method of Vapnik and Vladimir \cite{vapnik1963pattern}, which characterizes a set of data by separating it from the origin.
The \gls{nu-svm} method adds the kernel method, allowing non-linear decision functions, and incorporates `softness' by the $\nu$-parameter.
Whereas the method in \cite{vapnik1963pattern} focuses on two-class problems, the \gls{nu-svm} method introduces the method the one-class problems.

@@

The inner workings of \glspl{oc-svm} are discussed in detail in Section \ref{sec:one_class_svm}.

% \section{Change-detection by Support Vector Machines}\label{svm}
% *** REMOVE THIS SECTION? ***

@@

% The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.
% One-class SVMs are used to describe the current data under consideration, by assuming all data points are from the same class \cite{tax2001one}.
% To cope with possible errors or outliers a soft-margin is applied \cite{cortes1995support}.
% The class is described by a spherical boundary around the data with center $c$ and radius $r$, such that the volume is minimized.
% New data points are consecutively mapped from the input space to the feature space.
% The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.
% *** NOTE: using the in/out position of a new datapoint is different from the approach in this thesis, by only using the model properties. ***
% Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
% \begin{equation}
%   \mathrm{Min}\ r^2
% \end{equation}
% \begin{equation}
%   \mathrm{Subject\ to} : \|\mathbf{x}_i - \mathbf{c}\|^2 \le r^2\ \forall i,\ \mathbf{x}_i : i \mathrm{th\ data\ point}
% \end{equation}

@@

% *** Give visual explanation with circles and in- and out-side new data points ***

% *** Make clear what approach we use (model properties) and create link/bridge to next Chapter (3), which discusses SVM in detail. ***

>>>>

Many proposals in the field of remote sensing and \gls{har} make use of \glspl{svm} as a (supervised) model learning method.
An elaborate overview of applications is given in \cite{mountrakis2011support}.
In this section we review methods of change detection based on the applications of \glspl{svm} as model construction method.
A number of proposals have been made using this method.
Sch{\"o}lkopf \etal \cite{scholkopf1999support} applies Vapnik's principle, to never solve a problem which is more general than the one that one is actually interested in, to novelty detection.
In the case of novelty detection, they argue there is no need for density estimation of the distribution.
Simple algorithms estimate the density by considering how many data points fall in a region of interest.
The \gls{nu-svm} method instead starts with the number of data points that should be within the region and estimates a region with that desired property.
It builds on the method of Vapnik and Vladimir \cite{vapnik1963pattern}, which characterizes a set of data by separating it from the origin.
The \gls{nu-svm} method adds the kernel method, allowing non-linear decision functions, and incorporates `softness' by the $\nu$-parameter.
Whereas the method in \cite{vapnik1963pattern} focuses on two-class problems, the \gls{nu-svm} method introduces the method the one-class problems.

@@

The inner workings of \glspl{oc-svm} are discussed in detail in Section~\ref{sec:one_class_svm}.

\TODO{Add notion of `supervised' vs `unsupervised' learning. \gls{oc-svm} is unsupervised.}

% \section{Change-detection by Support Vector Machines}\label{svm}
% \TODO{REMOVE THIS SECTION?}

@@

% The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.
% One-class SVMs are used to describe the current data under consideration, by assuming all data points are from the same class \cite{tax2001one}.
% To cope with possible errors or outliers a soft-margin is applied \cite{cortes1995support}.
% The class is described by a spherical boundary around the data with center $c$ and radius $r$, such that the volume is minimized.
% New data points are consecutively mapped from the input space to the feature space.
% The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.
% \TODO{NOTE: using the in/out position of a new datapoint is different from the approach in this thesis, by only using the model properties.}
% Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
% \begin{equation}
%   \mathrm{Min}\ r^2
% \end{equation}
% \begin{equation}
%   \mathrm{Subject\ to} : \|\mathbf{x}_i - \mathbf{c}\|^2 \le r^2\ \forall i,\ \mathbf{x}_i : i \mathrm{th\ data\ point}
% \end{equation}

@@

% \TODO{Give visual explanation with circles and in- and out-side new data points}

% \TODO{Make clear what approach we use (model properties) and create link/bridge to next Chapter (3), which discusses SVM in detail.}

### 2013-11-01 Chapters/Chapter2/temporal_segmentation.tex
Prefix refs with non-breaking space and capitalize refs

% ``An online algorithm for segmenting time series'' \cite{keogh2001online}. See \ref{sec:appendix-C-online-keogh}. 538, 2001 \\
% ``Segmenting time series: A survey and novel approach'' \cite{keogh2004segmenting}. 242, 2004 \\

@@

% ``Joint segmentation of multivariate time series with hidden process regression for human activity recognition'' \cite{chamroukhi2013joint}. 0, 2013. See \ref{sec:appendix-C-joint-segmentation}. \\

@@

Section \ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.

>>>>

% ``An online algorithm for segmenting time series'' \cite{keogh2001online}. See~\ref{sec:appendix-C-online-keogh}. 538, 2001 \\
% ``Segmenting time series: A survey and novel approach'' \cite{keogh2004segmenting}. 242, 2004 \\

@@

% ``Joint segmentation of multivariate time series with hidden process regression for human activity recognition'' \cite{chamroukhi2013joint}. 0, 2013. See~\ref{sec:appendix-C-joint-segmentation}. \\

@@

Section~\ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.

### 2013-11-01 Chapters/Chapter3.tex
Prefix refs with non-breaking space and capitalize refs

\label{Chapter3} % For referencing the chapter elsewhere, use \ref{Chapter3}

@@

This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in Chapter \ref{Chapter4}.
The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
It transforms the traditional problem to change detection for times series data.
In Section \ref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in section \ref{sec:one_class_svm}.
For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in \ref{sec:sensor_data_characteristics}.
Finally a theoretic set of quality metrics is discussed in Section \ref{sec:quality_metrics}.

>>>>

\label{Chapter3} % For referencing the chapter elsewhere, use~\ref{Chapter3}

@@

This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in Chapter~\ref{Chapter4}.
The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
It transforms the traditional problem to change detection for times series data.
In Section~\ref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in Section~\ref{sec:one_class_svm}.
For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in~\ref{sec:sensor_data_characteristics}.
Finally a theoretic set of quality metrics is discussed in Section~\ref{sec:quality_metrics}.

### 2013-11-01 Chapters/Chapter3/change_detection_time_series.tex
Prefix refs with non-breaking space and capitalize refs

\TODO{Note: the following block is copied from \ref{sec:method_change_indication}} \\
The problem of change point detection can be formulated using different type of models, as discussed in \ref{subsec:segmentation}.
The methods by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} use the following formulation for change detection, which we will also use for our problem formulation.
The algorithm searches for \emph{sudden} changes in the time series data.
In other words, slowly changing properties in the data are not considered to be changes.
Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kl} divergence $D(p^2||p^1)$.
The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.

The definition of change point being \emph{sudden} changes in the time series data is in line with the search of changes in activities.
Since we are only interested in different activities (which are represented by sudden changes), slight changes within an activity are not of interest.
Section \ref{sec:sensor_data_characteristics} discusses the representation of activities and changes between them in the data in more detail.

>>>>

\TODO{Note: the following block is copied from~\ref{sec:method_change_indication}} \\
The problem of change point detection can be formulated using different type of models, as discussed in~\ref{subsec:segmentation}.
The methods by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} use the following formulation for change detection, which we will also use for our problem formulation.
The algorithm searches for \emph{sudden} changes in the time series data.
In other words, slowly changing properties in the data are not considered to be changes.
Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kl} divergence $D(p^2||p^1)$.
The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.

The definition of change point being \emph{sudden} changes in the time series data is in line with the search of changes in activities.
Since we are only interested in different activities (which are represented by sudden changes), slight changes within an activity are not of interest.
Section~\ref{sec:sensor_data_characteristics} discusses the representation of activities and changes between them in the data in more detail.
